### 2016-11-10 Paper_v1.tex
Nov7
Nov9

The ensemble of models captured the time evolution of cell free CAT biosynthesis (Fig.~\ref{fig:BothCarbon} - \ref{fig:Amino}).
Glucose was exhausted with 3 hr [FILL ME IN].
The ensemble also captured the energy species dynamics, particularly the overall energy total (Fig.~\ref{fig:BothCarbon}, top) and the totals by base .
The ensemble and the best-fit set also predict some of the amino acid measurements, while failing to predict others (Fig.~\ref{fig:Amino}).
the central carbon metabolism, including glucose uptake, CAT production, and the dynamics of the organic acid intermediates .
Allosteric control is important to the dynamics of the organic acid intermediates, as without it several of the measurements are not captured by the ensemble or the best-fit set (Fig.~\ref{fig:BothCarbon}, bottom).
This is likely due to a structural deficiency in the model; in some cases, the consumption of an amino acid through CAT synthesis is not enough to explain the decrease shown in the data, and there are no other reactions that consume it. Thus, a more comprehensive biological description is needed to fully explain amino acid dynamics.

@@

\subsection*{Sensitivity analysis}
We conducted a local sensitivity analysis to determine which of the kinetic and control parameters affected model performance.
We calculated performance as area under the CAT curve, which was directly related to CAT synthesis rate, as the culture time was fixed and no CAT degradation was modeled.
We randomly chose 180 sets of the 18,000 sets in the ensemble and defined these as our sub-ensemble; for each set in the sub-ensemble, we varied the rate constant and saturation constants of each metabolic flux and measured the resulting change in CAT production to estimate the sensitivity of model performance to that parameter.
We did the same for the control parameters, both the order (Hill coefficient) and gain (related to the the dissociation constant).
This allowed us to estimate the relative importance of the kinetic and control parameters to CAT production across the ensemble.
Of the rate constants, those with the highest positive sensitivities were CAT synthesis, GTP synthesis, GMP synthesis, glutamine synthesis, and aspartate synthesis (Fig.~\ref{fig:LocalSensitivity}, top).
This is explained by GTP and the amino acids being reactants for CAT synthesis.
Also, GMP synthesis increases the total amount of guanosine, allowing for more GTP production.
The rate constants with the largest negative sensitivities were GTP degradation, arginine synthesis, and UMP synthesis.
While GTP degradation is obvious, the others can be explained in that they consume ATP as well as several amino acids, all of which are reactants for CAT synthesis.
Of the saturation constants, the reverse is seen: the largest positive sensitivities are those associated with arginine synthesis, while the largest negative sensitivities are those associated with CAT synthesis, GTP synthesis, and GMP synthesis (Fig.~\ref{fig:LocalSensitivity}, middle).
This is because an increase in saturation constant causes a decrease in the corresponding reaction rate.
The control parameters were seen to be the least significant and the most uncertain (Fig.~\ref{fig:LocalSensitivity}, bottom).
Only two had a small standard error across the ensemble, relative to the ensemble mean sensitivity: the gain and order for pyruvate acting as an inhibitor on the pdh reaction.
This could be because pdh consumes pyruvate and diverts carbon away from the pathways that ultimately contribute to CAT production.
Taken together with the lack of change in glucose uptake and CAT production when control is removed (Fig.~\ref{fig:BothCarbon}), this suggests that allosteric control is not the limiting factor to CAT production.

We conducted a global sensitivity analysis on the parameters that could be controlled experimentally: the initial conditions of glucose, oxygen, amino acids, and enzymes.
We used the variance-based method of Sobol, and the same objective function of area under the CAT curve.
Using a parameter set of relatively good fit against data, we defined parameter bounds and generated a Sobol sequence of 111,600 parameter values that fit within those bounds.
We then calculated the total-order sensitivity and confidence interval for each of the experimentally controllable initial conditions.
As the sensitivities were total-order, they were guaranteed to be non-negative.
The largest sensitivities belonged to the initial conditions of the CAT macromolecular synthesis machinery, GTP synthase, GTP degradation, some amino acids such as phenylalanine, proline, and leucine, and some amino acid synthases (Fig.~\ref{fig:GlobalSensitivity}).
This is all explained by GTP and amino acids being reactants for CAT synthesis.
While some of the amino acids and amino acid synthases were among the highest in sensitivity, theirs were also very uncertain relative to the CAT macromolecular synthesis machinery and GTP synthase.
The initial conditions of glucose and oxygen were among the least important according to the global sensitivity analysis, suggesting that the model predicts that CAT production can be sustained by consuming initial stores or can be powered by other means.

@@

\subsection*{Local sensitivity analysis}
We determined the reactions most important to protein production by computing the local sensitivity of CAT concentration to each rate constant in the network.
Each sensitivity index was formulated as:
\begin{equation}\label{eqn:local-sensitivity}
\begin{split}
    S_{ij}=norm(CAT(p_i,p_j,t)-CAT(\alpha*p_i,\alpha*p_j,t))\qquad{i,j=1,2,\hdots\mathcal{P}}\\
\end{split}
\end{equation}
where $S_{ij}$ denotes the sensitivity of CAT production to the $i$th and $j$th parameters, $CAT(p_i,p_j,t)$ denotes CAT concentration as a function of time and the $i$th and $j$th parameters, $\alpha$ denotes the perturbation factor, equal to 1.01, and $\mathcal{P}$ denotes the number of rate constants, equal to 163.
In calculating the pairwise sensitivities, each parameter was perturbed by 1\%;

@@

% Figures and captions go here ...
\begin{figure}[ht]
\centering
\includegraphics[width=0.75\textwidth]{./Figures/CAT_prod_yield.pdf}
\caption{Sequence specific flux balance analysis of CAT production and yield. A. 95\% confidence interval of the ensemble (light blue region) for CAT concentration versus time. B. Theoretical maximum carbon yield of CAT calcualted by ssFBA for four different cases: unconstrained (black), limited oxidative phosphorylation activity (gray), constrained by transcription and translation rates (light grey), and constrained by experimental data (white). Error bars represent standard deviation of the ensemble.   }
\label{fig:CATProdYield}
\end{figure}

@@

\begin{figure}[ht]
\centering
%\includegraphics[width=1.00\textwidth]{./Figures/BothCarbon.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/Carbon.pdf}
\includegraphics[width=0.05\textwidth]{./Figures/null.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/CarbonNoControl.pdf}
\caption{Central carbon metabolism in the presence (top) and absence (bottom) of allosteric control, including glucose (substrate), CAT (product), and intermediates, as well as total concentration of energy species. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 100 sets.}
\label{fig:BothCarbon}
\end{figure}

\begin{figure}[ht]
\centering
%\includegraphics[width=1.00\textwidth]{./Figures/BothCarbon.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/Carbon.png}
\includegraphics[width=0.05\textwidth]{./Figures/null.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/CarbonNoControl.png}
\caption{Central carbon metabolism in the presence (top) and absence (bottom) of allosteric control, including glucose (substrate), CAT (product), and intermediates, as well as total concentration of energy species. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 100 sets.}
\label{fig:BothCarbon_png}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Energy.png}
\caption{Energy species and energy totals by base in the presence of allosteric control. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 100 sets.}
\label{fig:Energy_png}
\end{figure}

%\begin{table}[ht]
%\centering
%\includegraphics[width=0.94\textwidth]{./Figures/AIC.pdf}
%\caption{}
%\label{fig:AIC}
%\end{table}

\begin{figure}[ht]
\centering
\includegraphics[width=0.94\textwidth]{./Figures/ClusterCAT161.pdf}
\caption{Normalized first-order and pairwise sensitivities of CAT production to rate constants. Dark pixels indicate the most sensitive indices. Dendrogram indicates high responders (dark gray background), medium responders (medium gray background), and low responders (light gray background).}
\label{fig:ClusterCAT}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=0.94\textwidth]{./Figures/ClusterCAT161.png}
\caption{Normalized first-order and pairwise sensitivities of CAT production to rate constants. Dark pixels indicate the most sensitive indices. Dendrogram indicates high responders (dark gray background), medium responders (medium gray background), and low responders (light gray background).}
\label{fig:ClusterCAT_png}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/ClusterSystem.pdf}
\caption{Normalized first-order and pairwise sensitivities of system state to rate constants. Dark pixels indicate the most sensitive indices. Dendrogram indicates high responders (dark gray background), medium responders (medium gray background), and low responders (light gray background).}
\label{fig:ClusterSystem}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/AA_bounded_bar.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/AA_bounded_scatter.pdf}
\caption{Total-order global sensitivities of protein export flux to amino acid import flux upper bounds. Each upper bound was varied from 0 to 0.5 mM/hr.}
\label{fig:AAs_bounded}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/AA_unbounded_bar.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/AA_unbounded_scatter.pdf}
\caption{Total-order global sensitivities of protein export flux to amino acid import flux upper bounds. Each upper bound was varied from 0 to 0.5 mM/hr, except the upper bound with respect to which sensitivity was calculated; this was varied from 0 to 10 mM/hr.}
\label{fig:AAs_unbounded}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/C_Secretion.pdf}
\caption{Total-order global sensitivities of protein export flux to glutamic acid, oxygen and glucose import flux upper bounds, and translation initiation rate upper bound. Lactate, ethanol, acetate, and formate export fluxes were held constant at 0.1 mM/hr in the control case. One of these export fluxes was held constant at 10 mM/hr, while the others were held constant at 0.1 mM/hr, in each of the other four cases.}
\label{fig:C_Secretion.pdf}
\end{figure}

>>>>

The ensemble of models captured the time evolution of cell free CAT biosynthesis (Fig.~\ref{fig:CarbonBoth} - \ref{fig:Amino}).
Glucose was exhausted with 3 hr [FILL ME IN].
The ensemble also captured the energy species dynamics, particularly the overall energy total (Fig.~\ref{fig:CarbonBoth}, top) and the totals by base .
The ensemble and the best-fit set also predict some of the amino acid measurements, while failing to predict others (Fig.~\ref{fig:Amino}).
the central carbon metabolism, including glucose uptake, CAT production, and the dynamics of the organic acid intermediates .
Allosteric control is important to the dynamics of the organic acid intermediates, as without it several of the measurements are not captured by the ensemble or the best-fit set (Fig.~\ref{fig:CarbonBoth}, bottom).
This is likely due to a structural deficiency in the model; in some cases, the consumption of an amino acid through CAT synthesis is not enough to explain the decrease shown in the data, and there are no other reactions that consume it. Thus, a more comprehensive biological description is needed to fully explain amino acid dynamics.

\subsection*{Sensitivity analysis}
We performed a local sensitivity analysis to determine the network reactions with the greatest effect on protein production and overall system state.
CAT production was msot sensitive to the CAT synthesis reaction, oxidative phosphorylation activity, and alanine synthesis, as well as various reactions in glycolysis, the TCA cycle, and amino acid synthesis and degradation.

%We conducted a local sensitivity analysis to determine which of the kinetic and control parameters affected model performance.
%We calculated performance as area under the CAT curve, which was directly related to CAT synthesis rate, as the culture time was fixed and no CAT degradation was modeled.
%We randomly chose 180 sets of the 18,000 sets in the ensemble and defined these as our sub-ensemble; for each set in the sub-ensemble, we varied the rate constant and saturation constants of each metabolic flux and measured the resulting change in CAT production to estimate the sensitivity of model performance to that parameter.
%We did the same for the control parameters, both the order (Hill coefficient) and gain (related to the the dissociation constant).
%This allowed us to estimate the relative importance of the kinetic and control parameters to CAT production across the ensemble.
%Of the rate constants, those with the highest positive sensitivities were CAT synthesis, GTP synthesis, GMP synthesis, glutamine synthesis, and aspartate synthesis (Fig.~\ref{fig:LocalSensitivity}, top).
%This is explained by GTP and the amino acids being reactants for CAT synthesis.
%Also, GMP synthesis increases the total amount of guanosine, allowing for more GTP production.
%The rate constants with the largest negative sensitivities were GTP degradation, arginine synthesis, and UMP synthesis.
%While GTP degradation is obvious, the others can be explained in that they consume ATP as well as several amino acids, all of which are reactants for CAT synthesis.
%Of the saturation constants, the reverse is seen: the largest positive sensitivities are those associated with arginine synthesis, while the largest negative sensitivities are those associated with CAT synthesis, GTP synthesis, and GMP synthesis (Fig.~\ref{fig:LocalSensitivity}, middle).
%This is because an increase in saturation constant causes a decrease in the corresponding reaction rate.
%The control parameters were seen to be the least significant and the most uncertain (Fig.~\ref{fig:LocalSensitivity}, bottom).
%Only two had a small standard error across the ensemble, relative to the ensemble mean sensitivity: the gain and order for pyruvate acting as an inhibitor on the pdh reaction.
%This could be because pdh consumes pyruvate and diverts carbon away from the pathways that ultimately contribute to CAT production.
%Taken together with the lack of change in glucose uptake and CAT production when control is removed (Fig.~\ref{fig:CarbonBoth}), this suggests that allosteric control is not the limiting factor to CAT production.

%We conducted a global sensitivity analysis on the parameters that could be controlled experimentally: the initial conditions of glucose, oxygen, amino acids, and enzymes.
%We used the variance-based method of Sobol, and the same objective function of area under the CAT curve.
%Using a parameter set of relatively good fit against data, we defined parameter bounds and generated a Sobol sequence of 111,600 parameter values that fit within those bounds.
%We then calculated the total-order sensitivity and confidence interval for each of the experimentally controllable initial conditions.
%As the sensitivities were total-order, they were guaranteed to be non-negative.
%The largest sensitivities belonged to the initial conditions of the CAT macromolecular synthesis machinery, GTP synthase, GTP degradation, some amino acids such as phenylalanine, proline, and leucine, and some amino acid synthases (Fig.~\ref{fig:GlobalSensitivity}).
%This is all explained by GTP and amino acids being reactants for CAT synthesis.
%While some of the amino acids and amino acid synthases were among the highest in sensitivity, theirs were also very uncertain relative to the CAT macromolecular synthesis machinery and GTP synthase.
%The initial conditions of glucose and oxygen were among the least important according to the global sensitivity analysis, suggesting that the model predicts that CAT production can be sustained by consuming initial stores or can be powered by other means.

@@

@@

\subsection*{Sensitivity analysis}
We determined the reactions most important to protein production by computing the local sensitivity of CAT concentration to each rate constant in the network.
Each sensitivity index was formulated as:
\begin{equation}\label{eqn:local-sensitivity}
\begin{split}
    S_{ij}=norm(CAT(p_i,p_j,t)-CAT(\alpha*p_i,\alpha*p_j,t))\qquad{i,j=1,2,\hdots\mathcal{P}}\\
\end{split}
\end{equation}
where $S_{ij}$ denotes the sensitivity of CAT production to the $i$th and $j$th parameters, $CAT(p_i,p_j,t)$ denotes CAT concentration as a function of time and the $i$th and $j$th parameters, $\alpha$ denotes the perturbation factor, equal to 1.01, and $\mathcal{P}$ denotes the number of rate constants, equal to 163.
In calculating the pairwise sensitivities, each parameter was perturbed by 1\%; first-order sensitivities ($i$ = $j$) were subject to two 1\% perturbations, equivalent to a perturbation of 2.01\%.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/CarbonBoth.pdf}
\caption{Central carbon metabolism in the presence (top) and absence (bottom) of allosteric control, including glucose (substrate), CAT (product), and intermediates, as well as total concentration of energy species. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 100 sets.}
\label{fig:CarbonBoth}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/BoxPlot-KCFM.pdf}
\caption{Log of cost function across 37 datasets for data-trained ensemble (blue) and randomly generated ensemble (red, gray background). Median (bars), interquartile range (boxes), range excluding outliers (dashed lines), and outliers (circles) for each dataset. Median across all datasets (large bar overlaid).}
\label{fig:BoxPlot}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth,trim=0 0 263 0,clip]{./Figures/ClusterCAT.pdf}
\caption{Normalized first-order and pairwise sensitivities of CAT production to rate constants.}
\label{fig:ClusterCAT}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.15\textwidth,trim=0 0 263 0,clip]{./Figures/ClusterSystem.pdf}
\caption{Normalized first-order and pairwise sensitivities of system state to rate constants. A (most sensitive parameters): hexokinase run by GTP, and lactate dehydrogenase forward reaction. B: }
\label{fig:ClusterSystem}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=0.75\textwidth]{./Figures/CAT_prod_yield.pdf}
\caption{Sequence specific flux balance analysis of CAT production and yield. A. 95\% confidence interval of the ensemble (light blue region) for CAT concentration versus time. B. Theoretical maximum carbon yield of CAT calcualted by ssFBA for four different cases: unconstrained (black), limited oxidative phosphorylation activity (gray), constrained by transcription and translation rates (light grey), and constrained by experimental data (white). Error bars represent standard deviation of the ensemble.}
\label{fig:CATProdYield}
\end{figure}

@@
